Columnar VCF for Biobank scale datasets using Zarr

### 2024-04-04 paper.tex

\subsection{Case study: Genomics England 100,000 genomes}
In this section we demonstrate the utility of vcfzarr on a large human dataset.
[DETAILS ABOUT GeL DATA]
We converted the data to the intermediate columnar format 
which took X hours with a cluster of Y workers using the 
distributed explode commands. When then converted to Zarr using 
SOMETHING which took X hours using [RESOURCES]. In total, this gave 
X TiB of data, spread across Y thousand files, a compression of 
X over the original VCFs.
The detailed breakdown of per-column compression ratios is 
shown in Table~\ref{tab-genomics-england-data}.

>>>>

\subsection{Case study: Genomics England 100,000 genomes}
In this section we demonstrate the utility of vcfzarr on a large human dataset.
Genomics England’s multi-sample VCF dataset (aggV2) is an 
aggregate of 78,195 gVCFs from rare disease and cancer participants 
recruited as part of the 100,000 Genomes Project~\cite{turnbull2018100}. 
aggV2 comprises approximately 722 million annotated single-nucleotide 
variants (SNVs) and small indels split into 1,371 roughly equal chunks and 
totalling 165.3 Tib of data after \texttt{bgzip} compression. 
The dataset is used for a variety of research purposes, ranging from 
genome-wide association studies (GWAS)~\cite{kousathanas2022whole} and 
imputation~\cite{shi2023genomics} to 
simple queries involving single gene 
regions~\cite{leggatt2023genotype}~\cite{lam2023repeat}.
We converted the data to the intermediate columnar format 
which took X hours with a cluster of Y workers using the 
distributed explode commands. When then converted to Zarr using 
SOMETHING which took X hours using [RESOURCES]. In total, this gave 
X TiB of data, spread across Y thousand files, a compression of 
X over the original VCFs.
The detailed breakdown of per-column compression ratios is 
shown in Table~\ref{tab-genomics-england-data}.
